\chapter{Measurement Problems}

In our everyday world, measuring mostly doesn't pose any difficulties.
Let's take baking a cake as an easy example.
If you want to bake a cake, you'll have a recipe with all required ingredients in the correct amounts.
A good kitchen has all the tools needed to measure the amounts: a measuring cup for the milk, a scale for the flour or simply your mind to count the eggs.
In this example, it is very intuitively what must be measured.
The given amounts of ingredients are essential. 
Too little or too much could ruin the cake.
Of course all ingredients are needed to get the cake right, none can be left out.
It's also clear what measurement tool is needed for which ingredient.
One would be very lost if they were to measure the circumference of the eggs or weigh ingredients that don't go in the cake at all.
Because this is a simple example, we easily realize that unnecessary measurements are absurd.

Measuring turns more into a problem when we are confronted with more abstract cases.
An example of this is the measuring of code quality.
Now the tools required to measure aren't as obvious anymore, it's also not as easy to know what exactly to measure.
One of the first questions that come to mind is why even measure code quality?
Measuring code quality is needed to verify its readability, maintainability and reusability.
To control these three aspects, we have a variety of variables we can measure: the complexity of the code, the number of defects in the code or whether standards have been followed or not.
Lastly, we need to know how we can measure those variables. 
We can count the comment lines, lines of code or calculate defect density.
These are only very few options of why, what and how to measure code quality.
In this more abstract example, where things aren't as clear anymore, it's much easier to commit unnecessary or even wrong measures.
This is why we need methods to prevent "measuring the circumference of the eggs".

In the context of business are also countless variables that can be measured: financial success, customer satisfaction, production cost, brand value, product quality.
To only name a few.
Here we face the same diffuclties about why, what and how to measure as in the example the the code.
Darrell Mann introduces a variety of tools to overcome problems with measuring in the business world \cite{darrell2004hands}.
In the following pages we'll analyze Mann's tools and discuss the methods for solving measurement problems.
















